Rapid, (semi-?)Supervised Generation of Novel Percussive Sounds With A Tractable Virtual Synthesizer

### 2020-01-03 paper.tex
Updates from Overleaf

\end{enumerate}
Automatic programming of virtual synthesizers has also been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines \cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al. \cite{yee2018automatic} used LSTM and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain, even with random search. Since this work more focused on pads and textures rather than drums, feature matching in this also appears to not be concerned with the envelope of the sounds, but rather the frequency content within arbitrary time windows. Yet another recent, impressive work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio \cite{esling2019universal}. As stated before, our work presents a rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation. 
\begin{table}[h!]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |c|c|c| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & tone type\\
IsNoise & boolean & whether to generate noise\\
Sample Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start<1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cuttoff & 0-20000 & -\\
LP filter Cuttoff & HP-20000 & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer Submodule Parameters}
\label{table:1}
\end{table}

\section{Implementation}
\label{impl}

@@

\caption{Model working well on a test set}

@@

\section{Novel Generations}
\label{gens}
\subsection{The Goal}
 Given a category, our aim is to use random generations to guide an algorithm towards novel generation of drums for that category. We evaluate each method by measuring how many samples of categories \textit{kick,snare} and \textit{hat} can be produced in under a minute. Finally we show results of a survey conducted to measure the quality of these samples. 
\subsection{General Methods}
\subsubsection{Random Generation}
To establish a benchmark, we seek to establish the likelihood of generating samples of each category if we generate programs for various stack sizes by random selection of parameters. \\
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-a}
\includegraphics[width=0.48\linewidth]{example-image-a}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:rand-graphs}
\end{figure}
\subsubsection{Multivariate Normal}
Some success and extremely fast and scalable once the parameter space is calculated. A well known issue with this apporach is that the parameter space for virtual synthesizers are often neither differenciable nor gaussian \cite{esling2019universal} (a couple more papers also talked about this, I have to find them). Currently our parameters are array indices mapping to different values for Synth components. This makes the assumption that our discrete parameter space is normally distributed or differentiable in-appropriate and prone to errors. \\
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-b}
\includegraphics[width=0.48\linewidth]{example-image-b}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:multi-graphs}
\end{figure}

@@

\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-c}
\includegraphics[width=0.48\linewidth]{example-image-c}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:evo-graphs}
\end{figure}
\subsubsection{Survey of Generated Drums}
\label{survey}
\colorbox{green!=30}{Finalize ear,generate,make survey.} The survey will be sent out. I know a few producers who are willing to do this.
3 algorithms, 3 categories, 10 samples for each algorithm and category pair, 2 seconds to listen to each sample-> 3 minutes for listening\\
With those numbers I imagine survey will be around 5-6 minutes long.
\subsubsection{Contrasting Ear Scores With Survey Results}
\colorbox{green!=30}{pending section \ref{survey}}
\section{Code and Database}
\colorbox{green!=30}{One last thing to do, will take a 1 day}\\
\colorbox{blue!=30}{Should this section be at the start of the paper?Appendix?}\\
talk about  categories, sources, data processing. 
Code for the data processing is written, I need to download the no-copyright sounds, run the code and host the results. 
Give link for people to download it, show where to find the scripts for database expansion.
\section{Summary and Conclusions}
\colorbox{green!=30}{Points I want to make:}
\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item With the right implementation of DSP methods to create a synth, Genetic search is viable for finding  desired programs and parameters.
    \item When possible, using DSP methods instead of NNs is preferable in many ways(speed,computation,hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\colorbox{blue!=30}{Need to think about this more}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
    \item Can our approach be used for VSTs?
\end{itemize}
%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.
\section{Ethical Standards}
To ensure objectivity and transparency in research and to ensure that accepted principles of ethical and professional conduct have been followed, authors should include a section “Ethical Standards” before the References, including (if relevant): information regarding sources of funding, potential conflicts of interest (financial or non-financial), informed consent if the research involved human participants, statement on welfare of animals if the research involved animals.

>>>>

\end{enumerate}
Automatic programming of virtual synthesizers has also been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines \cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al. \cite{yee2018automatic} used LSTM and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain, even with random search. Since this work more focused on pads and textures rather than drums, feature matching in this also appears to not be concerned with the envelope of the sounds, but rather the frequency content within arbitrary time windows. Yet another recent, impressive work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio \cite{esling2019universal}. As stated before, our work presents a rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation. 
\begin{table}[h!]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |c|c|c| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & tone type\\
IsNoise & boolean & whether to generate noise\\
Sample Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start<1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cuttoff & 0-20000 & -\\
LP filter Cuttoff & HP-20000 & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer Submodule Parameters}
\label{table:1}
\end{table}
\subsection{Data And Project Replication}
\colorbox{green!=30}{One last thing to do, will take a 1 day}\\
\colorbox{green!=30}{need to finalize database}\\
\colorbox{blue!=30}{Should this section be at the start of the paper?The End? Appendix?}\\
talk about  categories, sources, data processing. 
Code for the data processing is written, I need to download the no-copyright sounds, run the code and host the results. 
Give link for people to download it, show where to find the scripts for database expansion.
\section{Implementation}
\label{impl}

@@

\caption{confusion matrix of the model working well on a test set. \colorbox{green!=30}{will be updated with new ear.}}

@@

\section{Novel Generations}
\label{gens}
\subsection{The Goal}
 Given a category, our aim is to use random generations to guide an algorithm towards novel generation of drums for that category. We evaluate each method by measuring how many samples of categories \textit{kick,snare} and \textit{hat} can be produced in under a minute. Finally we show results of a survey conducted to measure the quality of these samples. 
\subsection{General Methods}
\colorbox{green!=30}{The next 3 sections will be modified heavily after the ear is updated.}
\subsubsection{Random Generation}
To establish a benchmark, we seek to establish the likelihood of generating samples of each category if we generate programs for various stack sizes by random selection of parameters. We have observed a bias in generations as indicated in figure \ref{gens}.\\
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-a}
\includegraphics[width=0.48\linewidth]{example-image-a}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:rand-graphs}
\end{figure}
\subsubsection{Multivariate Distribution Sampling}
Some success and extremely fast and scalable once the parameter space is calculated. A well known issue with this apporach is that the parameter space for virtual synthesizers are often neither differenciable nor gaussian \cite{esling2019universal} (a couple more papers also talked about this, I have to find them). Currently our parameters are array indices mapping to different values for Synth components. This makes the assumption that our discrete parameter space is normally distributed or differentiable in-appropriate and prone to errors. \\
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-b}
\includegraphics[width=0.48\linewidth]{example-image-b}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:multi-graphs}
\end{figure}

@@

\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-c}
\includegraphics[width=0.48\linewidth]{example-image-c}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:evo-graphs}
\end{figure}
\subsubsection{Survey of Generated Drums}
\label{survey}
\colorbox{green!=30}{Finalize ear,generate,make survey.} The survey will be sent out. I know a few producers who are willing to do this.
3 algorithms, 3 categories, 10 samples for each algorithm and category pair, 2 seconds to listen to each sample-> 3 minutes for listening\\
With those numbers I imagine survey will be around 5-6 minutes long.
\subsubsection{Contrasting Ear Scores With Survey Results}
\colorbox{green!=30}{pending section \ref{survey}}

\section{Summary and Conclusions}
\colorbox{green!=30}{Points I want to make:}
\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item Reiterate the importance of ear.
    \item There are potential benefits (e.g: pedagogical, software engineering, efficiency, data search, generality of solution, etc.) to separate some audio generation tasks into multiple parts, rather than throwing everything at some neural net. 
    \item With the right implementation of DSP methods to create a synth, Genetic search (viewed as too slow by previous work) is viable for finding  desired programs and parameters.
    \item Neural nets are function approximators. You don't need them if the actual function is understandable and simple...(but is our task simple enough? how do I prove it?)
    \item When possible, using DSP methods instead of NNs is preferable in many ways(speed,computation,hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\colorbox{blue!=30}{Need to think about this more}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
    \item Can our approach be used for VSTs?
\end{itemize}
%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.
\section{Ethical Standards}
To ensure objectivity and transparency in research and to ensure that accepted principles of ethical and professional conduct have been followed, authors should include a section “Ethical Standards” before the References, including (if relevant): information regarding sources of funding, potential conflicts of interest (financial or non-financial), informed consent if the research involved human participants, statement on welfare of animals if the research involved animals.
